---
title: "EDA"
authors:
- Kevin Sheng
- Arushi Agarwal
- Peter Kim
- Nathan Dang
output:
  html_notebook:
    code_folding: show
    highlight: haddock
    theme: lumen
    toc: yes
    toc_depth: 4
    toc_float: yes
---

```{r setup, include = FALSE}
# echo = FALSE because there's so much boilerplate code jesus
knitr::opts_chunk$set(echo = FALSE, cache = T, fig.width = 6, fig.height = 3)
options(scipen = 0, digits = 3)  # controls base R output
if (!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(tidyverse, data.table, usmap)
```

```{r initialize crap, include = FALSE}
data <- fread("data/final_data.csv")
```

Let's start off with some rudimentary things.

```{r rudimentary stuff}
data %>%
        select(-c("fips", "fips_state", "fips_county")) %>%
        summary()
```

# Population

For some weird reason, `pop` doesn't take a normal distribution.

```{r}
data %>% ggplot() + geom_histogram(aes(x = pop), bins = 100)
```

Even if we only take the counties from the far, *far* left of the plot, it's still heavily skewed to the left. \
This shows that the huge majority of counties in the US have a population of somehwere between 5k and 100k.

```{r}
# to prevent the super big counties from making the rest really darn small
pop_lim <- 2e5
lim_data <- data %>% filter(pop < pop_lim)
lim_data %>% ggplot() + geom_histogram(aes(x = pop), bins = 150)
```

Let's try plotting population against some response variables. Maybe larger cities have some things in common.

```{r}
points <- c("property crime", "violent crime")
lim_data %>%
        ggplot(aes(x = pop)) +
        geom_jitter(aes(y = property_crime_rate, colour = points[1])) +
        geom_jitter(aes(y = violent_crime_rate, colour = points[2])) +
        scale_color_manual("type", breaks = points, values = c("blue", "red")) +
        xlab("population") +
        ylab("crime per 100k people") +
        labs(title = "crime rates vs. population")

points <- c("unemployment rate", "poverty rate")
lim_data %>%
        ggplot(aes(x = pop)) +
        geom_jitter(aes(y = unemployment_rate, colour = points[1])) +
        geom_jitter(aes(y = poverty, colour = points[2])) +
        scale_color_manual("type", breaks = points, values = c("purple", "red")) +
        xlab("population") +
        ylab("percentage") +
        labs(title = "percentage stuff vs. population")

points <- c("median household income", "per capita income")
lim_data %>%
        ggplot(aes(x = pop)) +
        geom_jitter(aes(y = median_household_income, colour = points[1])) +
        geom_jitter(aes(y = per_capita_income, colour = points[2])) +
        scale_color_manual("type", breaks = points, values = c("green", "yellow")) +
        xlab("population") +
        ylab("income ($)") +
        labs(title = "income vs. population")
```

Well, those weren't the results I was hoping for. \
By the way, the county with that really high property crime rate is Bibb County in Georgia.

Now let's try plotting the same things against `hs_grad`:

```{r}
points <- c("property crime", "violent crime")
data %>%
        ggplot(aes(x = hs_grad)) +
        geom_jitter(aes(y = property_crime_rate, colour = points[1])) +
        geom_jitter(aes(y = violent_crime_rate, colour = points[2])) +
        scale_color_manual("type", breaks = points, values = c("blue", "red")) +
        xlab("high school grad rate") +
        ylab("crime per 100k people") +
        labs(title = "crime rates vs. high school grad rate")

points <- c("unemployment rate", "poverty rate")
data %>%
        ggplot(aes(x = hs_grad)) +
        geom_jitter(aes(y = unemployment_rate, colour = points[1])) +
        geom_jitter(aes(y = poverty, colour = points[2])) +
        scale_color_manual("type", breaks = points, values = c("purple", "red")) +
        xlab("high school grad rate") +
        ylab("percentage") +
        labs(title = "percentage stuff vs. high school grad rate")

points <- c("median household income", "per capita income")
data %>%
        ggplot(aes(x = hs_grad)) +
        geom_jitter(aes(y = median_household_income, colour = points[1])) +
        geom_jitter(aes(y = per_capita_income, colour = points[2])) +
        scale_color_manual("type", breaks = points, values = c("green", "yellow")) +
        xlab("high school grad rate") +
        ylab("income ($)") +
        labs(title = "income vs. high school grad rate")
```

Much better! \
It is a bit weird that `hs_grad` seems to correlate pretty strongly with poverty, but not with unemployment. \
It also seems to kind of relate to property crime, but not really to violent crime.

What if we graphed property crime against violent crime?

```{r}
data %>%
        ggplot() +
        geom_jitter(aes(x = violent_crime_rate, y = property_crime_rate)) +
        labs("property crime rate vs. violent crime rate")
```

It seems there's a **lot** of outliers- there does seem to be a *slight* positive correlation.

Out of curiosity, I also graphed the output variables against `bachelors`.

```{r}
points <- c("property crime", "violent crime")
data %>%
        ggplot(aes(x = bachelors)) +
        geom_jitter(aes(y = property_crime_rate, colour = points[1])) +
        geom_jitter(aes(y = violent_crime_rate, colour = points[2])) +
        scale_color_manual("type", breaks = points, values = c("blue", "red")) +
        xlab("bachelor rate") +
        ylab("crime per 100k people") +
        labs(title = "crime rates vs. bachelor rate")

points <- c("unemployment rate", "poverty rate")
data %>%
        ggplot(aes(x = bachelors)) +
        geom_jitter(aes(y = unemployment_rate, colour = points[1])) +
        geom_jitter(aes(y = poverty, colour = points[2])) +
        scale_color_manual("type", breaks = points, values = c("purple", "red")) +
        xlab("bachelor rate") +
        ylab("percentage") +
        labs(title = "percentage stuff vs. bachelor rate")

points <- c("median household income", "per capita income")
data %>%
        ggplot(aes(x = bachelors)) +
        geom_jitter(aes(y = median_household_income, colour = points[1])) +
        geom_jitter(aes(y = per_capita_income, colour = points[2])) +
        scale_color_manual("type", breaks = points, values = c("green", "yellow")) +
        xlab("bachelor rate") +
        ylab("income ($)") +
        labs(title = "income vs. bachelor rate")
```

Nothing really seems to correlate with `crime_rate`... \
`bachelor` does seem to have a negative relationship with `unemployment` though, that's pretty good!

I also tried graphing `broadband` and `computer` against `median_household_income` and got this:

```{r}
points <- c("broadband access %", "computer access %")
data %>%
        ggplot(aes(x = median_household_income)) +
        geom_jitter(aes(y = broadband, colour = points[1])) +
        geom_jitter(aes(y = computer, colour = points[2])) +
        scale_color_manual("type", breaks = points, values = c("cyan", "blue")) +
        xlab("median household income") +
        ylab("% of pop") +
        labs(title = "tech access measures vs. median household income")
```
A very positive correlation, but it's painfully obvious when you realize that richer families will have more purchasing power for tech products and the like. \
Also, because the correlation that `computer` and `broradband` have are so similar, I'm just gonna use `computer`

Moving onto ethnicity:

```{r race pivots}
races <- data %>%
        select(
                black, native, asian, pac_isl,
                hispanic, white_not_hispanic,
                other_single_race, two_plus_races
        ) %>%
        pivot_longer(cols = everything(), names_to = "race", values_to = "pct")
races$race <- as.factor(races$race)
```

these are the box plots of the percentages of each race in each community
```{r plotting races}
races %>%
  ggplot() +
  geom_boxplot(aes(x = race, y = pct, fill = race))
```

There's a lot of hispanic-majority and black-majority communities, but those communities are in the minority.

```{r}
# regions = "counties", for county level summary
plot_usmap(regions = "counties",
           data = data,
           values = "hs_grad",
           exclude = c("Hawaii", "Alaska"),
           color = NA) +
  scale_fill_gradient(low = "white", high = "red",
                      name = "High school grad percentage",
                      label = scales::comma) +
  labs(title = "Counties high school graduation rate",
       subtitle = "Continental US Counties") +
  theme(legend.position = "right")

plot_usmap(regions = "counties",
           data = data,
           values = "bachelors",
           exclude = c("Hawaii", "Alaska"),
           color = NA) +
        scale_fill_gradient(low = "white", high = "red",
                            name = "Bachelor percentage",
                            label = scales::comma) +
        labs(title = "Counties bachelor degree rate",
             subtitle = "Continental US Counties") +
        theme(legend.position = "right")
```
Most counties has a relatively similar (high) value for `hs_grad`, except for some counties down south.
