---
title: "Final Project"
authors:
- Kevin Sheng
- Arushi Agarwal
- Peter Kim
- Nathan Dang
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = T, fig.width=6, fig.height=3)
options(scipen = 0, digits = 3)  # controls base R output
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(tidyverse, data.table)
```

* [unemployment data](https://www.bls.gov/lau/)
* [basically everything else](https://www.openintro.org/data/index.php?data=county_complete)
* [crime data](https://ucr.fbi.gov/crime-in-the-u.s/2017/crime-in-the-u.s.-2017)
* [FIPS table & stuff](https://github.com/kjhealy/fips-codes/blob/master/state_and_county_fips_master.csv)

note that the csv files we used were slightly modified to remove/include some useless/useful data

```{r initialization}
data <- fread("data/final_data.csv")
```

output variables (for each county):
* crime rate
* income
* med. income

interesting input variables
* high school graduate

control input variables (added a buncha controls):
* population
* race percentages
  * black
  * asian
  * white
  * hispanic

eda is [here](eda.rmd)
