---
title: "Final Project"
authors:
- Kevin Sheng
- Arushi Agarwal
- Peter Kim
- Nathan Dang
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = T, fig.width=6, fig.height=3)
options(scipen = 0, digits = 3)  # controls base R output
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(tidyverse, data.table)
```

* [unemployment data](https://www.bls.gov/lau/)
* [basically everything else](https://www.openintro.org/data/index.php?data=county_complete)
* [crime data](https://ucr.fbi.gov/crime-in-the-u.s/2017/crime-in-the-u.s.-2017)
* [FIPS table & stuff](https://github.com/kjhealy/fips-codes/blob/master/state_and_county_fips_master.csv)

note that the csv files we used were slightly modified to remove/include some useless/useful data

```{r initialization}
data <- fread("data/final_data.csv")
```

some variables are in absolutes, some are in percentages, don't know if this'll screw up the relations

* output variables (for each county):
  * crimes per 100k ppl (`violent_crime_rate` & `property_crime_rate`)
  * unemployment (`unemployment_rate`)
  * med. income/per capita income, who knows (`median_household_income` or `per_capita_income`)
  * poverty rate (`poverty`)
* interesting input variables
  * high school graduate % (`hs_grad`)
  * some college education % (`some_college`)
  * at least bachelor's % (`bachelors`)
* control input variables (yeah there's a lot of them):
  * population (`pop`)
  * demographics (`black`, `native`, etc.)
  * veteran % (`veterans`)
  * broadband & computer access % (`broadband` & `computer`)
  * household information (`households`, `persons_per_household`)

eda is [here](eda.rmd)
